Characterizing Stellar Streams by ML

### 2023-08-11 src/tex/ms.tex
[showyourwork] overleaf sync

\newcommand{\mrm}[1]{\mathrm{#1}}
\newcommand{\mbs}[1]{\boldsymbol{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mfk}[1]{\mathfrak{#1}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\unit}[1]{[\text{#1}]}

\newcommand{\nth}[1]{{#1}_{\mrm{n}}}
\newcommand{\fth}[1]{{#1}_{\mrm{f}}}
\newcommand{\qth}[1]{{#1}_{\mrm{q}}}

\newcommand{\pdf}{P}
\newcommand{\cdf}{\Phi}

\newcommand{\Exp}[1]{e^{#1}}
\newcommand{\Erf}[1]{\mathrm{erf}\left(#1\right)}

\newcommand{\sigobs}{{\sigma_*}}
\newcommand{\JN}[1]{{\textcolor{red}{JN: #1}}}

\NewDocumentCommand \dif {o m} {
	\IfNoValueTF{#1}
		{ \mathrm{d}{#2} }
		{ \mathrm{d}^{#1}{\!#2} }
}

@@

% Author list
\author{Nathaniel Starkman}
\author{Jacob Nibauer}
\author{Jo Bovy}
\author{Jeremy Webb}

@@

    [Content Here]

    \

    Paper Map:
    \begin{enumerate}
        \item Intro
        \item Method \begin{enumerate}
            \item Fig 1: Overall explainer figure that illustrates the method with Neural networks, and norm flow for CMD
            \item Fig 2: Bayesian graphical model
            \item Test On Synthetic Stream:
            \item Some sort of plot demonstrating successful recovery
        \end{enumerate}
        \item Application / Results \begin{enumerate}
            \item Test case GD-1: \begin{enumerate}
                \item CMD norm flow background
                \item Fig: Stream phi1 phi2 top panel and histograms of each dimension below GD-1
                \item Some plot in phi1-phi2 that shows density variations
            \end{enumerate}
            \item Test case Pal-5 \begin{enumerate}
                \item CMD norm flow background
                \item Fig: Stream phi1 phi2 top panel and histograms of each dimension below Pal 5
                \item Some plot in phi1-phi2 that shows density variations
            \end{enumerate}
        \end{enumerate}
        \item Conclusion
        \item Appendix
    \end{enumerate}

@@

    Stellar streams

@@

        \begin{figure}
        \centering
        \includegraphics[scale=0.5]{figures-not-syw/DAG_StreamDensity.pdf}
        \caption{Fill in.}
        \label{fig: DAG}
        \end{figure}

        We model the stream and background as functions continuous in some affine parameter. In principle this parameter can be anything. Though it might seem the best parameter is the stream's arc length, the most common, easiest, and practical parameter is the longitude -- $w_{\cdot, j\equiv0} \equiv \phi_1$ -- in some stream-oriented reference frame \citep[for details see][section 2]{2023MNRAS.522.5022S}.

        We model the likelihood for a single star as
        \begin{align} \label{eq:general_model}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
                &= \sum_{q} \ f_q(w_{i, 0}) \pdf(\nth{\mbs{w}}|\mbs{\theta},q_i\!=\!q) \pdf(\nth{\mbs{m}}|\mbs{\theta},q_i\!=\!q)
        \end{align}
        where we have assumed that $\nth{\mbs{w}}$ and $\nth{\mbs{m}}$ are conditionally independent, and that the mixture weights $f$ are normalized s.t. $\sum_q f(w_{i,0}) = 1$. We discuss the mixture weights further in \autoref{sub:mixture_weight}

@@

        \begin{multline}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta}) 
            = \\\phantom{+} \ \quad f \quad\ P^{(S)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(S)}(\nth{\mbs{m}}|\mbs{\theta})  \\
            \quad + (1-f) P^{(B)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(B)}(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{multline}

        The probablity of a being a stream member is then simply
        \begin{equation}\label{eq:membership_prob}
            \pdf\left(S | \mbs{w}_n, \mbs{m}_n, \mbs{\theta} \right) = \frac{f P^{(S)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(S)}(\nth{\mbs{m}}|\mbs{\theta}) }{ \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})}.
        \end{equation}
    Hereafter, when we refer to ``membership probability," Eq.~\ref{eq:membership_prob} provides the formal definition. 
    %\newpage

    \subsection{Determination of the Mixture Weight} \label{sub:mixture_weight}
        The mixture weight $f_q(w_{i,0})$ as used in Eq.~\ref{eq:general_model} is characterized as the fraction of stars belonging to the mixture component $q$ over an increment $[w_{i,0}, w_{i,0} + \dif{w_{i,0}}]$. Thus, the mixture weight encodes information about the linear density of the stream, which has been shown to depend on substructure in the galaxy (CITE-x). The dependence of the mixture weight on the linear density is captured by the expression
        \begin{equation}\label{eq:fraction_param}
            f_q\left(w_{i,0}\right) = \frac{N_q\left(\phi_1\right)}{N_{\rm tot}\left(\phi_1\right)},
        \end{equation}
        where $N_q\left(\phi_1\right)$ is the number of stars in component $q$ within an interval $[\phi_1, \phi_1 + \dif{\phi_1}]$, and 
        $N_{\rm tot}$ is the total number of stars in the same band (including those in component $q$). The linear  density for component $q$ is then entirely specified by the function $N_q(\phi_1)$.

@@

        In particular, for a given field the distribution of $\phi_1$ coordinates is specified by the set $\{\phi_{1,i}\}_i$. A normalizng flow is trained to this dataset, and will satisfy $\int d\phi_1 P_{\phi_i}(\phi_1) = 1$ by construction. In order to re-weight the normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of stars in the field considered. We can then obtain the linear stream density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q \left[N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) \right] .
        \end{equation}

@@

            The astrometric track is modeled with a neural network, encompassing each phase space dimension, parameterized as a function of $\phi_1$. The neural network output for a single stream component is the mean location of the stream track in each dimension, $\mbs{\mu}_w(\phi_1 | q_i)$, and its associated covariance matrix $\mbs{\Sigma}_w(\phi_1 | q_i)$. The probability for the data is then
            \begin{equation}
                \pdf^{(S)}(\nth{\mbs{w}} | \mbs{\theta}, \phi_{1,i}, q_i) = \mcal{N}\big(\nth{\mbs{w}} | \mbs{\mu}_w\left(\phi_{1,i} | q_i \right), \mbs{\Sigma}_w\left(\phi_{1,i} | q_i \right) \big),
            \end{equation}
            where $\mbs{\theta}$ is the set of neural network parameters that define $\mbs{\mu}_w$ and $\mbs{\Sigma}_w$.
            In the simplest case, we take $\mbs{\Sigma}_w$ to be diagonal, consisting of both intrinsic dispersion for each dimension of $\mbs{w}$ and the data uncertainties added in quadrature.

            In practice, we work with truncated Gaussians, where the truncation is set by the size of the field in each astrometric dimension. This ensures that the probability density is zero wherever there is no data. Details of the truncated Gaussian are included in \autoref{app:distributions}.

@@

            We now outline the off-stream astrometric model. We take a similar approach to the on-stream astrometric model discussed in \S\ref{ssub:astrometric_model_on_stream}, though we typically do not characterize the backgrounds with a Gaussian distribution. Instead, we utilize a range of distributions (i.e., exponential, skew-normal) discussed in \autoref{app:distributions}. The parameters of the user-specified background distributions are themselves neural network outputs, such that any $\phi_1$ ``slice" of the background should be characterized by some analytic density, but the evolution of the backgrounds with $\phi_1$ can be complex.

            Let $\mbs{\theta}_{B,j}(\phi_1)$ represent the neural network which outputs a vector of parameters characterizing the background distribution at a given $\phi_1$ for astrometric dimension $j$. The number of outputs for the background model is equal to $\Sigma_{j>0} \mathrm{dim}\left(\mbs{\theta}_{B,j}\right)$ (where $j=0$ corresponds to $\phi_1$, which we do not model). At a given $\phi_1$, the probability density for the astrometric dimension $j>0$ is $\pdf^{(B)}_j\left(\mbs{w}_{n,j} | \mbs{\theta}, \mbs{\theta}_{B,j}, \phi_1 \right)$. We treat each background component as independent, allowing us to write the likelihood as a product over the astrometric dimensions:
            \begin{equation}
                \pdf^{(B)}\left(\mbs{w}_n | \mbs{\theta}, \phi_1 \right) = \prod_{j>0} \pdf_j^{(B)}\left(\mbs{w}_{n,j} | \mbs{\theta}, \mbs{\theta}_{B,j}, \phi_1 \right).
            \end{equation}

@@

            \todo[inline]{TODO:}
            \begin{enumerate}
                \item Not including isochrone parameters, e.g. age and metallicity
                \item integrating over $\gamma$ the model assumes a PDMF, e.g. Kroupa
            \end{enumerate}

            By integrating over $\gamma$ the model is not sensitive to

@@

@@

        \begin{figure*}
            \script{mock/plot/results.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/mock/results.pdf}
            \caption{Caption: TO-DO need to show this with a ground truth (e.g. show true weight in top panel).}
            \label{fig:mock_data_result}
        \end{figure*}

\section{Model Implementation}

    \subsection{Network Architecture}

        \begin{enumerate}
            \item A network per model component, e.g. one for the weights, one for the stream track and width, one for the background distribution.
            \item MLP for the components - lin / tanh + dropout
            \item Normalizing flow - specific pieces
        \end{enumerate}

        \begin{table}
            \centering
            \caption{MLP structure}
            \begin{tabular}{@{}ccc@{}}
            \toprule
            Layer & Shape & Param \# \\
            \midrule
            Linear & (1, 3) & 1  \\
            Tanh & (2, 3) & 1  \\
            Dropout & (2, 3) & 0  \\
            \bottomrule
            \end{tabular}
        \end{table}

    \subsection{Astrometric Stream}

        \subsubsection{Track Region Prior} \label{sub:track_region_prior}

@@

            % The total log-probability over all the

            % \begin{equation}
            %     \ln\mcal{L}\left(\{(\nth{\mbs{w}},\nth{\mbs{m}})\} | \mbs{\theta}\right) = \sum_i  \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta}).
            % \end{equation}

>>>>

% Change the section names
\renewcommand{\sectionautorefname}{\S\!}%
\renewcommand{\subsectionautorefname}{\S\!\!}%
\renewcommand{\subsubsectionautorefname}{\S\!}%
\renewcommand{\equationautorefname}{Eq.}

\newcommand{\package}[1]{\textsc{#1}}
\newcommand{\stream}[1]{#1}
\newcommand{\dataarchive}[1]{\texttt{#1}}

% Math fonts
\newcommand{\mrm}[1]{\mathrm{#1}}
\newcommand{\mbs}[1]{\boldsymbol{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mfk}[1]{\mathfrak{#1}}
\newcommand{\mcal}[1]{\mathcal{#1}}

% Math symbols
\newcommand{\Exp}[1]{e^{#1}}
\newcommand{\Erf}[1]{\mathrm{erf}\left(#1\right)}
\NewDocumentCommand \dif {o m} {
	\IfNoValueTF{#1}
		{ \mathrm{d}{#2} }
		{ \mathrm{d}^{#1}{\!#2} }
}

% Statistics
\newcommand{\pdf}{\mcal{P}}
\newcommand{\cdf}{\mfk{P}}
\newcommand{\likelihood}{\mcal{L}}
\newcommand{\prior}{\mcal{\pi}}
\newcommand{\posterior}{p}

% Labels
\newcommand{\nth}[1]{{#1}_{\mrm{n}}}  % measurement
\newcommand{\fth}[1]{{#1}_{\mrm{f}}}  % feature
\newcommand{\qth}[1]{{#1}_{\mrm{q}}}  % model index
\newcommand{\unit}[1]{[\text{#1}]}

\newcommand{\cmp}[2]{#2^{(#1)}}
\newcommand{\smallcomponent}[2]{#2^{\scriptscriptstyle (#1)}}
\newcommand{\astroM}[1]{\smallcomponent{w}{#1}}
\newcommand{\photoM}[1]{\smallcomponent{m}{#1}}
\newcommand{\chemoM}[1]{\smallcomponent{z}{#1}}

\newcommand{\sigobs}{{\sigma_*}}

\newcommand{\JN}[1]{{\textcolor{red}{JN: #1}}}

@@

% Author list
\author{Nathaniel Starkman \& Jacob Nibauer}
\author{Jo Bovy}
\author{Jeremy Webb}
\author{Ana Bonaca}

@@

    \todo[inline]{Rephrase ---}

    \begin{quotation}

        A stellar stream is the association of stars tidally stripped from a common
        progenitor system that orbits in an underlying galactic gravitational field
        \citep{Johnston1998, HelmiWhite1999}. Stream progenitors are gravitationally
        bound systems, such as satellite dwarf galaxies or star clusters
        \citep{Odenkirchen2001, Majewski2003}. In the absence of external interactions
        these systems are stable, though still capable of mass loss driven by internal
        processes \citep{Hills1975, HeggieHut2003}, e.g. relaxation and binary
        interactions. Considering external interactions, these progenitor systems may
        be disrupted: completely, as in the merger events that formed the Galactic
        stellar halo \citep{Lynden-Bell1967, Searle1978}; or partially, such as tidal
        disruption forming a stellar stream.

        In the approximation of a smooth galactic gravitational field, the
        progenitor's extent is limited to the tidal radius. Through processes internal
        and external, progenitor stars can achieve sufficient energy to bring them
        beyond the tidal boundary, thus ``escaping'' into the galactic field. Escape
        primarily occurs at both an inner and outer saddle point of the combined
        potential -- e.g. the L2 and L3 Lagrange points for a spherical galactic field
        \citep{Ross1997, Fukushige2000, BinneyTremaine2008}. Generally, the initial
        phase space position of the escaped star is very similar to that of the
        progenitor: the configuration position is the tidal boundary and the star has
        a low peculiar velocity. Since both the escaped star and its progenitor orbit
        in the same underlying galactic potential and have similar phase space
        positions their orbits are similar \citep{BinneyTremaine2008}. Given
        sufficient time the phase space separation will increase and an escaped star
        will lead or trail the position of the progenitor, depending on the location
        from which it escaped. Therefore, a collection of tidally stripped stars will
        form extended structures, aka tails or streams, that are few to thousands of
        parsecs in length \citep{Johnston1998, HelmiWhite1999, Bovy2014}.

        To be a coherent, observable structure the peculiar velocity dispersion of
        escaped stars relative to the progenitor must be relatively low, but does vary
        from system to system. Stars in kinematically hot streams (e.g. from
        progenitors with larger kinematic dispersions) have more dissimilar orbits
        from each other and the progenitor than do stars in cold streams, with smaller
        dispersions \citep{Johnston1998, Hendel2015}. The orbit of a star depends on
        its phase space position and the gravitational potential; therefore measuring
        a star's position and velocity and observing its orbit reveals the local
        potential, which is normally dominated by the Galaxy (primarily the dark
        matter halo at large distances) \citep{Gibbons2014}. However, an orbit is
        impractical to observe on human timescales. Cold streams offer a workaround.
        Like measuring one star at many points in one orbit over an extended period,
        streams are many stars measured at different points along approximately one
        orbit at the same time. Therefore stellar streams are one of the most
        promising means to study the Galactic gravitational potential: the stellar
        components as well as the dark matter \citep{Johnston2016}.

        On the observational front there has been a large concerted effort in the
        field to find new stellar streams \citep[for an extensive list of known
        Galactic streams see][]{Mateu2022}, extend the detected extents of known
        streams \citep[e.g. for \stream{Palomar 5} alone:][]{Rockosi2002, Ibata2017,
        Grillmair2006, Carlberg2012, StarkmanEtAl2019}, and increase the general
        purity of stream catalogs.

    \end{quotation}

    \todo[inline]{--- Rephrase}

    \paragraph{Review of detection methods}

        Identifying streams of globular of clusters is a challenging task. Streams are very low signal-to-noise structures spread over comparatively large regions of the sky. They live in a vast background of field stars.

        Historically streams were detected by matched-filter techniques
        % Add citations
        You took an isochrone and shifted it by a distance modulus until a stream stood out in longitude and latitude as an extended overdensity structure.

        With the advent of Gaia, we gained access to detailed astrometric data. For most streams this

        poses a challenging task due to their intricate and diffuse nature
        from large-scale surveys is challenging due to the complex and overlapping distributions of stars in different regions of the sky; disentangling members from the vast background of field stars

        \begin{enumerate}
            \item Streamfinder \citep{2018MNRAS.477.4063M}. This method requires the potential
            \item VIA MACHINAE \citep{2022MNRAS.509.5992S}
            \item Kiyan and Adrian
            \item Matched Filter
            \item Gaia stuff: Adrian and Ana and Sara et al
            \item Uniform Modeling of 12 Stellar stream \citep{PatrickEtAl2022}. Only on phi1, phi2, and CMD. Their CMD method is more sophisticated. Worth discussing.
        \end{enumerate}

    Review of available data:
    \begin{enumerate}
        \item Gaia. Streams in Gaia have on-sky positions and proper motions, and parallaxes, though for most streams the parallax errors are prohibitively large. Likewise, most streams in Gaia do not have good radial velocity measurements. The photometry is generally poor, generally necessitating cross-matching with other datasets.
        \item S5. This is a much smaller dataset. It has deeper photometry and radial velocity measurements.
    \end{enumerate}

    \paragraph{Paper organization}

        The paper is organized as follows.
        In \autoref{sec:method} we introduce our method for characterizing stream tracks and computing stellar membership.
        We discuss the data in \autoref{sub:framing_the_data} and build a probabilistic framework to model the stream and background in \autoref{sub:likelihood_setup}-\autoref{sub:priors}.
        In \autoref{sec:results} we apply our method,
        % TODO: rephrase the mock data
        first to simulated data (\autoref{sub:mock_data}) to demonstrate the model with a known ground truth dataset,
        then to the stellar streams \stream{GD-1} (\autoref{sub:gd1}) and \stream{Pal\,5} (\autoref{sub:pal5}).
        We compare our approach to characterizing stellar streams to other methods in \autoref{sec:comparison}.
        We conclude in \autoref{sec:conclusions}, discussing the results of this work and future directions.

@@

    We now introduce our method for characterizing stellar stream tracks and computing stream membership likelihoods from N-D phase-space measurements.
    % TODO: rephrase following sentence
    We develop a flexible Bayesian model framework, using Mixture Density Networks, that allows arbitrary numbers of phase-space dimensions to be incorporated in the model and permits missing phase-space observations.

    % \subsection*{}  Conceptual Description

        Below, we offer a concise overview of our approach, followed by a more in-depth technical explanation in subsequent sections.

        \paragraph{one orbit and a single population}

            Stellar streams consist of stars that are tidally disrupted from a progenitor system.
            To a first approximation, the stars occupy a single orbit, that of their progenitor.
            Additionally, as stars formed in the progenitor they retain its photometric and chemical characteristics: which for globular clusters means following an isochrone.
            In this simplistic picture, astrometrically characterizing a stream requires only finding it's orbit. Member stars fall directly on the orbit and the isochrone, making membership likelihood determination trivial.

            % Observational errors complicate this picture. Instead of a certain membership calculation, we must build a probabilistic model, assigning a likelihood that an observed star falls along the stream's orbit.

        \paragraph{similar orbits and a somewhat noisy population}

            Occupying a single orbit implies that streams remain a point in action space, and having a common origin would thus not become extended structure. This is not true.
            Differences in stream stars' orbits from the progenitor's arise from various factors including internal dynamics of the progenitor and the host Galactic potential.

            An intrinsic width, in configuration and kinematic space.

            Centered on a track, not an orbit.

            This defines an astrometric phase-space hyper-tube \cite{2018MNRAS.477.4063M}, with a distribution radial to the track.

            Instead of characterizing the stream by its orbit, we characterize it by it's track and the distribution around that track.
            Computing membership likelihoods now requires fitting the track and distribution. This cannot be
            done without then also fitting the field of stars in which the stream lives, since the stream 
            and field cannot be clearly separated.

        \paragraph{gaps orbits and a somewhat noisy population}

            This is still a simplistic picture of streams.

            Perturbations to the Galactic potential, in the form of time evolution and non-axisymmetry, as well as substructure can disrupt the stellar stream. 
            Even most perturbations don't ruin the idea of the phase-space hypertube. They either drop the mixture weight (rephrase, not yet introduced) to 0 or change the width of the hypertube.

        % TODO: significantly expand on this

        \paragraph{Modeling approach}

            We take a Bayesian approach modeling the stream and background distribution.

            Mixture model are a general formalism for modelling arbitrary conditional density functions.

            In this case we want to model the density function, conditioned on the location along the track.

            However, previously we discussed how stream will vary along the track, with changes in the width and possible gaps.
            Also observational effects, e.g the distance gradient changes the on-sky width (should this be mentioned earlier?)

            We can take the parameters of the mixture model -- the mixing coefficients and component distribution parameters -- to be general (continuous) functions of the conditioning variable.
            We do this by modelling the parameters as This is achieved outputs of a feed-forward neural network that takes the conditioning variable as its input. This is called a Mixture Density Network (MDN)

    % subsection framing_the_data (end)

    \subsection{Framing the Data} \label{sub:framing_the_data}

        The hypertube idea is connected to orbits.
        One might think then that a variable like the phase-angle or the arc-length, or any variable that directly sweeps along the stream would be best to use as the conditioning variable.
        This is not the case for a number of reasons (see \cite{StarkmanEtAl2023} for some discussion on this topic), not least of which is that we do not have any of these quantities. We do not know the stream track, it's what we want to know.

        Instead we will approach this in the standard way: by parameterizing the stream along its the longitude -- $w_{\cdot, j\equiv0} \equiv \phi_1$ -- in a stream-oriented reference frame.
        $\phi_1$ becomes our conditioning variable.

        \begin{quotation}
            The stream is projected from its initial reference frame, e.g. ICRS
            \citep{ICRS1997}, into a rotated on-sky frame, with longitude $\phi_1$ and
            latitude $\phi_2$, such that the stream extends along $\phi_1$. In this frame
            the stream may be fit by a low-order polynomial as a function of $\phi_1$
            \citep[e.g.][]{Bonaca2020}. Similar, if more sophisticated, methods swap tools
            like univariate splines \citep[e.g.][]{Erkal2017, Bonaca2018} for the
            low-order polynomials. Using a coordinate dimension, here $\phi_1$, as the
            independent variable works for many streams, as $\phi_1$ sweeps monotonically
            along the stream. For example \stream{GD-1} is straight for over 80 degrees
            \citep{Webb2019, Price-Whelan2018}.
        \end{quotation}

        There are known to be issues with this...

        \begin{quotation}
            However, this is not true in general: a
            stream's path can become kinked by subhalo interactions, or curved on the
            celestial sphere in an observer's reference frame, or like the Magellanic
            stream wrap multiple times around the Milky Way \citep{Wannier1972}. When a
            stream track is not a function, methods fitting $\phi_2(\phi_1)$ fail.
        \end{quotation}

        We leave extending the model framework to phase-wrapped, doubled-backed paths, etc to a future work.

@@

        \begin{figure}
        \centering
        \includegraphics[scale=0.5]{static/DAG_StreamDensity.pdf}
        \caption{Fill in.}
        \label{fig: DAG}
        \end{figure}

        We model the likelihood for a single star as
        \begin{align} \label{eq:general_model}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
                &= \sum_{q} \ f_q(w_{i, 0}) \pdf(\nth{\mbs{w}}|\mbs{\theta},q_i\!=\!q) \pdf(\nth{\mbs{m}}|\mbs{\theta},q_i\!=\!q)
        \end{align}
        where we have assumed that $\nth{\mbs{w}}$ and $\nth{\mbs{m}}$ are
        conditionally independent, and that the mixture weights $f$ are
        normalized s.t. $\sum_q f(w_{i,0}) = 1$.  We discuss the mixture weights
        further in \autoref{sub:mixture_weight}.

@@

        \begin{align}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} P^{(S)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(S)}(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) P^{(B)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(B)}(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}

        The probablity of a being a stream member is then simply
        \begin{equation}\label{eq:membership_prob}
            \pdf\left(S | \mbs{w}_n, \mbs{m}_n, \mbs{\theta} \right) = \frac{f P^{(S)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(S)}(\nth{\mbs{m}}|\mbs{\theta}) }{ \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})}.
        \end{equation}

        Hereafter, when we refer to ``membership probability,"
        \autoref{eq:membership_prob} provides the formal definition.

        \paragraph{missing phase-space observations}

            Streams have a diversity of measurement coverage: several have full 6-D astrometrics as well as photometrics and chemical abundances available \citep[e.g.,][]{Koposov2019, 2020A&A...635L...3A, 2022ApJ...928...30L}, while most have partial coverage, lacking feature dimensions.
            % TODO: terminology ND-phase-space vs feature columns
            Most streams in \dataarchive{Gaia} {\color{red} [cite]}, for instance, have proper motions but not radial velocities nor reliable parallax measurements. Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many feature dimensions, however only for a select set of streams. Given the diversity of coverage it is important that the likelihood model can deal with missing data.

            The method of treatment of missing data depends on the cause for that data to be missing.
            Broadly, there are two categories of causes: randomness and systematics.
            Consider a dataset with all features measured but with a random subset of features masked.
            One possible approach is to attempt to impute the missing features, then feed the full-featured data to the likelihood model. Alternatively, the likelihood model might use only the present features, ignoring those missing. If however, features are missing for systematic reasons, then this approach can introduce systematic bias. Consider a star too far away to measure it's parallax distance or radial velocity. This is a systematic reason and this information should be incorporated into any model as a prior. Not doing so, the data imputation approach might predict too-close distances. Likewise, using only present features in the likelihood model misses that the star is likely very distant.

            {\color{red} Sentence explaining why we ignore features regardless.}
            We include a quality flag on our calculations, indicating the number of missing features.

            The per-star likelihood is modelled with a distribution with the same dimensionality as the data vector.
            For example, a star with a 6D astrometric measurement will be modelled with a full 6D distribution, while a star missing its radial velocity measurement, will be modelled with a 5D dimensionally-reduced form of the 6D distribution. Using dimensionally reduced distributions ensure that missing phase-space dimensions do not amplify nor suppress the likelihood.

    \subsection{Determination of the Mixture Weight} \label{sub:mixture_weight}
        The mixture weight $f_q(\phi_1)$ as used in \autoref{eq:general_model} is characterized as the fraction of stars belonging to the mixture component $q$ over an increment $[\phi_1, \phi_1 + \dif{\phi_1}]$. Thus, the mixture weight encodes information about the linear density of the stream, which has been shown to depend on substructure in the galaxy (CITE-x). The dependence of the mixture weight on the linear density is captured by the expression
        \begin{equation}\label{eq:fraction_param}
            f_q\left(w_{i,0}\right) = \frac{N_q\left(\phi_1\right)}{N_{\rm tot}\left(\phi_1\right)},
        \end{equation}
        where $N_q\left(\phi_1\right)$ is the number of stars in component $q$ within an interval $[\phi_1, \phi_1 + \dif{\phi_1}]$, and 
        $N_{\rm tot}$ is the total number of stars in the same band (including those in component $q$). The linear  density for component $q$ is then entirely specified by the function $N_q(\phi_1)$.

@@

        In particular, for a given field the distribution of $\phi_1$ coordinates is specified by the set $\{\nth{\phi_{1}}\}_f$. A normalizng flow is trained to this dataset, and will satisfy $\int d\phi_1 P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of stars in the field considered. We can then obtain the linear stream density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q \left[N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) \right] .
        \end{equation}

@@

            The astrometric track is modeled with a neural network, encompassing each phase space dimension, parameterized as a function of $\phi_1$. The neural network output for a single stream component is the mean location of the stream track in each dimension, $\astroM{\mbs{\mu}}(\phi_1)$, and its associated covariance matrix $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is then
            \begin{equation}
                \pdf^{(S)}(\nth{\mbs{w}} | \astroM{\mbs{\theta}}(\nth{\phi_{1}})) = \mcal{N}_F \big(\nth{\mbs{w}} | \astroM{\mbs{\mu}} \left(\nth{\phi_{1}} \right), \astroM{\mbs{\Sigma}} \left(\nth{\phi_{1}}\right) \big),
            \end{equation}
            where $\mbs{\theta}$ is the set of neural network parameters that define $\astroM{\mbs{\mu}}$ and $\astroM{\mbs{\Sigma}}$.
            In the simplest case, we take $\astroM{\mbs{\Sigma}}$ to be diagonal, consisting of both intrinsic dispersion for each dimension of $\mbs{w}$ and the data uncertainties added in quadrature.

            In practice, we work with truncated Gaussians, where the truncation is set by the size of the field in each astrometric dimension. Details of the truncated Gaussian are included in \autoref{app:sub:normal_distribution}. Truncating to the field ensures that the probability density is zero wherever there is no data. This has an added benefit of correctly normalizing the Gaussian distribution, which is not compact (has non-zero probability over all space) and would otherwise not integrate to one within the field.

            As a MDN, we implement the Gaussian distribution parameters with a multi-layered feed-forward neural network. The number of layers required depends on the maximum dimensionality of the Gaussian, and whether the neural network employs regularization techniques like dropout. We use alternating layers of linear and tanh units. Details of the network architecture are discussed in \autoref{sub:model_implementation}.

            In the presence of missing phase-space coordinate $f_k$ we reduce the dimensionality of the Gaussian to that of the data-vector. Namely, for $\mbs{I}_{\tilde{F}} = \mbs{I}_{F} \backslash f_k$
            \begin{align}
                \nth{\tilde{\mbf{w}}} &\mapsto \nth{\tilde{\mbf{w}}} = \{ w_{n,f}\}_{f\in \mbs{I}_{\tilde{F}}}, \nonumber\\
                \astroM{\mbs{\mu}} &\mapsto \astroM{\tilde{\mbs{\mu}}} = \{ \mu_{f}\}_{f\in\mbs{I}_{\tilde{F}}}, \nonumber\\
                \astroM{\mbs{\Sigma}} &\mapsto \astroM{\tilde{\mbs{\Sigma}}} = \{ \Sigma_{i,j}\}_{i,j \in \mbs{I}_{\tilde{F}}}, & \nonumber
                \intertext{such that,}
                \pdf^{(S)}(\nth{\mbs{w}} & | \astroM{\mbs{\theta}}(\nth{\phi_{1}}))
                    \mapsto \pdf^{(S)}(\nth{\tilde{\mbs{w}}} | \astroM{\tilde{\mbs{\theta}}} (\nth{\phi_{1}})).
            \end{align}

@@

            We now outline the off-stream astrometric model. We take a similar approach to the on-stream astrometric model discussed in \autoref{ssub:astrometric_model_on_stream}, though we typically do not characterize the backgrounds with a Gaussian distribution. Instead, we utilize a range of distributions (i.e., exponential, skew-normal) discussed in \autoref{app:distributions}. The parameters of the user-specified background distributions are themselves neural network outputs, such that any $\phi_1$ ``slice" of the background should be characterized by some analytic density, but the evolution of the backgrounds with $\phi_1$ can be complex. Details of the network architecture are discussed in \autoref{sub:model_implementation}.

            Let $\fth{\cmp{B}{\mbs{\theta}}}(\phi_1)$ represent the neural network which outputs a vector of parameters characterizing the background distribution at a given $\phi_1$ for astrometric dimension $f$. The number of outputs for the background model is equal to $\mathrm{dim}\left(\fth{\cmp{B}{\mbs{\theta}}}\right)$ (where $f=0$ corresponds to $\phi_1$, which we do not model). At a given $\phi_1$, the probability density for the astrometric dimension $f>0$ is $\fth{\pdf^{(B)}}\left(\mbs{w}_{n,f} | \fth{\cmp{B}{\mbs{\theta}}}(\phi_1) \right)$. We treat each background component as independent, allowing us to write the likelihood as a product over the astrometric dimensions:
            \begin{equation} \label{eq:astrometric_model_off_stream_probability}
                \pdf^{(B)}\left(\nth{\mbs{w}} | \mbs{\theta}(\phi_1) \right) = \prod_{f\in \mbs{I}_{\tilde{F}}} \pdf_f^{(B)}\left(\mbs{w}_{n,f} | \fth{\cmp{B}{\mbs{\theta}}}(\phi_1) \right).
            \end{equation}
            Where $\mbs{I}_{\tilde{F}}$ is the set of indices of the features, omitting missing phase space dimensions.

@@

            \todo[inline]{TODO:}
            \begin{enumerate}
                \item Not including isochrone parameters, e.g. age and metallicity
                \item integrating over $\gamma$ the model assumes a PDMF, e.g. Kroupa
                \item generic details of neural network, e.g Table 1
                \item introducing parallax
                \item how we reduce dimensionality.
            \end{enumerate}

@@

            \todo[inline]{TODO:}
            \begin{enumerate}
                \item Overview of the architecture
                \item How one would reduce the dimensionalty.
            \end{enumerate}

@@

        \begin{figure*}
            \script{mock/plot/results.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/mock/results.pdf}
            \caption{Caption: .}
            TODO:
            \begin{enumerate}
                \item Get rid of the dropout in the plot
                \item the linear density in the stream-fraction plot
            \end{enumerate}
            \label{fig:mock_data_result}
        \end{figure*}

    \subsection{Priors} \label{sub:priors}

        \subsubsection{Bounding Boxes}

        \subsubsection{Region Prior} \label{sub:track_region_prior}

            Region priors encourage the model to pass through a user-defined control region, while also remaining compatible with the data.

@@

    \subsection{Linking Coordinates: Combining the Distance Modulus with the Parallax Track}

        \todo[inline]{TODO:}
        \begin{enumerate}
            \item Discuss how to connect the distance modulus to the parallax.
            \item First-order error propagation to set width DM to width parallax.
        \end{enumerate}

    \subsection{Model Implementation, aka the code} \label{sub:model_implementation}
